### 2008-03-13 trunk/Papers/dl-gre-08/bisim.tex
some more references
Carlos: First ideas on the intro to bisim.tex.  Added some packages in dl-gre-08.

\section{Computing bisimulation classes} \label{sec:bisim}

>>>>

\newcommand{\gM}{\mathcal{M}}
\newcommand{\gL}{\mathcal{L}}

\section{Description Languages and Bisimulation} \label{sec:bisim}

In computational linguistics, the genereting referring expression problem (the GRE
problem) can be intuitively presented as follows: given some information about
a group of objects (e.g., grounded facts about individuals in an ongoing dialogue),
and a grammar, found a grammatically correct expression (usually a noun phrase) that uniquely identify a given object.

By taking a logic perspective, we can recast the GRE problem as an inference problem.
We can think that the information about the group of objects is provided in the form
of a model $\gM$ of a given logical language $\gL$.  Now, the inference problem
associated with the GRE problem will be to find a formula in $\gL$ such that it
exacly identify a given object $o$.  Formally, if $|\cdot|^\gM$ is the interpretation
function for $\gL$ which for each formula in $\gL$ and for each model $\gM$ returns
the set of elements in $\gM$ for which the formula $\varphi$ holds, then we can define
\medskip

\noindent
{\small
\begin{tabular}{l} \hline
\textsc{$\gL$-GRE Problem}\\ \hline
\ \ \ Input: A model $\gM$, an object $o$ in the domain of $\gM$\\
\ \ \ Output: a formula $\varphi \in \gL$ such that $|\varphi|^\gM = \{o\}$\\
\hspace*{1.4cm} (if such a formula exists).\\ \hline
\end{tabular}}
\medskip

As it can be seen from the definition, the problem depends on which particular
logical language $\gL$ we are interested in, and also (as the original GRE problem)
it might terminate with failure.  Depending on the model $\gM$, the chosen object $o$
and the expressive power of $\gL$ it can be the case that there is no formula of
$\gL$ that uniquely identifies $o$.

Consider the following example.  We have three objects related as indicated in
Figure~\ref{fig-car1}
\begin{figure}
\begin{center}
\includegraphics{pic.pdf}
\end{center}
\caption{Distinguishable/indistinguishable points.}\label{fig-car1}
\end{figure}

The points $o$ and $o'$ can easily be distinguished in first-order logic:
the formula $R(x,x)$ is true of $o$ and false in $o'$. The corresponding
referring expression for $o$ would be ``a point relating to itself''.
But if we can only use propositional and relational information (equivalently, our grammar does not contain the word ``itself'') then $o$ and $o'$ are indistinguishable. Both
are ``points which are related to points which are related to points\ldots''

\cite{dovier04:_effic_algor_for_comput_bisim_equiv}
